---
title: "Generating_Imaginary_Marketing_Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(lubridate)
library(dplyr)
library(ggplot2)
library(plotly)
```

## Generating Some example Marketing Data to Build and Test the Shiny App with.

First of all, let's generate a range of dates which will represent our existing historical data range:

```{r Dates}
date_seq <- seq(ymd("2017-01-01"),ymd("2017-12-31"),"days")
historic <- data.frame(date = date_seq)
historic$week_day <- lubridate::wday(historic$date, label=TRUE)
historic
```

Next up for this excercise we will make some assumptions about how budget is planned on this imaginary account.
Each Month, for the last 5 days, there will be a sale.
When a sale is on, spend will increase,

So let's plan spend...
We will have a constant spend, which fluctuates randomly from day to day.
When a sale is on, spend will increase by 50% or so, nd the spend will increase each day until the end of the sale.

```{r Spend}
#generate a base spend rate, which varies a random amount each day:


#List the dates where we will be in sale, as a function of the last 5 days of each month.

#function for last day of the month

#build the list


```




## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
